How Slow do my Transient Typechecks Go?

### 2019-08-17 paper.tex
Further edit pass through S5

\section{Individual Annotations}
\label{s-individual}
\begin{figure*}
	\input{graphs/pattern}
	\caption{Pairs of colour coded scatter and column graphs. The scatter graphs represent the performance of a sample of typing lattice. The column graphs show the performance of every configuration with only one type annotation. The scatter plots and columns graphs are colour coded based on whether a particular type annotation or two are present in the source code.}
	\label{f:pattern}		
\end{figure*}
We hypothesise that the tri/bi-modal graphs shown in figure~\ref{f:full}, are caused by individual type annotations: i.e. certain annotations have a significant effect on performance. In order to identify such type annotations, for each benchmark and type annotation, we manually inspected the scatter plots for the benchmark's typing lattice, coloured based on whether the chosen type annotation was present or not. This resulted in 1,775 graphs, however most of these showed no discernible pattern.

As an alternative, and less labour intensive approach to identifying potentially influential type annotations, for each type annotation we ran each benchmark with only that type annotation present. We used the same computer and number of iterations as for the typing lattice experiments, albeit these were done on different days.

@@

List is interesting as it demonstrates that the red and blue type annotations both cause a significant decrease in performance, which is even greater when both are present. Richards is particularly odd as it shows a performance decrease that appears to occur when the red type annotations is present, but not the blue one.

@@

The remaining benchmarks, which we have not shown in \ref{f:pattern}, have relatively flat column graphs and we could not identify any patterns in the typing lattice. The benchmark with the greatest difference in single type configurations relative to the untyped configuration is Storage, with a configuration only $7.9\%$ faster than the untyped baseline. However, though Storage (as shown in \ref{f:full}) appears to be bimodal like Towers, we were unable to find any relationships between its typing lattice and any individual type annotations.
%For example for Snake, the dots covered in red (corresponding to configurations with type annotation number 61) are noticeably higher than the grey ones (those without type annotation 61), thus it appears that the 61\textsuperscript{st} type annotation causes a noticeable slowdown, which is also exemplified by the higher than average red column. The graphs for CD (type \#107), Towers (type \#7) show similar patterns. The graph for Permute is similar, but it shows that type annotation 8 actually makes the benchmarks \emph{faster}.
%In some cases, more than two type annotations have an effect, for DeltaBlue and Go it appears that types 154/155, and 31/47 (respectively) all cause an increase in performance, but this is not cumulative.
%However with List, the performance identified by type annotations 17 and 18 is actually comulative

>>>>

\section{Individual Annotations}
\label{s-individual}
\begin{figure*}
	\input{graphs/pattern}
	\caption{Pairs of colour coded scatter and column graphs. The scatter graphs represent the performance of a sample of typing lattice. The column graphs show the performance of every configuration with only one type annotation. The scatter plots and columns graphs are colour coded based on whether a particular type annotation or two are present in the source code.}
	\label{f:pattern}		
\end{figure*}
We hypothesise that the tri/bi-modal graphs, shown in figure~\ref{f:full}, are caused by individual type annotations: that is, certain annotations have a significant effect on performance. In order to identify such type annotations, for each benchmark and type annotation we manually inspected the scatter plots for the benchmark's typing lattice, coloured based on whether the chosen type annotation was present or not. This resulted in 1,775 graphs, however most of these showed no discernible pattern.

As an alternative, and less labour-intensive approach to identifying potentially-influential type annotations, for each type annotation we ran each benchmark with \textit{only} that type annotation present. We used the same computer and number of iterations as for the typing lattice experiments, albeit these were done on different days.

@@

List is interesting as it demonstrates that the red and blue type annotations both cause a significant decrease in performance, which is even greater when both are present. Richards is particularly odd as it shows a performance decrease that appears to occur when the red type annotation is present, but not the blue one.

@@

The remaining benchmarks, which we have not shown in \ref{f:pattern}, have relatively flat column graphs and we could not identify any patterns in the typing lattice. The benchmark with the greatest difference in single-type configurations relative to the untyped configuration is Storage, with a configuration only $7.9\%$ faster than the untyped baseline. However, though Storage (as shown in \ref{f:full}) appears to be bimodal like Towers, we were unable to find any relationships between its typing lattice and any individual type annotations.
%For example for Snake, the dots covered in red (corresponding to configurations with type annotation number 61) are noticeably higher than the grey ones (those without type annotation 61), thus it appears that the 61\textsuperscript{st} type annotation causes a noticeable slowdown, which is also exemplified by the higher than average red column. The graphs for CD (type \#107), Towers (type \#7) show similar patterns. The graph for Permute is similar, but it shows that type annotation 8 actually makes the benchmarks \emph{faster}.
%In some cases, more than two type annotations have an effect, for DeltaBlue and Go it appears that types 154/155, and 31/47 (respectively) all cause an increase in performance, but this is not cumulative.
%However with List, the performance identified by type annotations 17 and 18 is actually comulative
